---
title: "Exoloratory Data Analysis of the Covid-19 Data Set"
output: github_document
bibliography: ../doc/covid19.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
library(tidyverse)
library(knitr)
library(broom)
library(datateachr)
library(infer)
library(cowplot)
theme_set(theme_minimal())
set.seed(2020)
```

## Summary of the data set

The data set used in this project is a collection of the COVID-19 data maintained by [Our World in Data](https://ourworldindata.org/coronavirus). It is updated daily and includes data on confirmed cases, deaths, hospitalizations and testing, as well as other variable of potential interests. This data has been collected, aggregated, and documented by Cameron Appel et al. [@owidcoronavirus]. Various origins have been sourced. Confirmed cases, deaths, hospitalizations and other related information are sourced from the European Centre for Disease Prevention and Control (ECDC). Testing related information are collected by the Our World in Data team from offical reports. Other sources include United Nations, World Bank and Global Burden of Disease. More information can be found [here](https://github.com/owid/covid-19-data/blob/master/public/data/owid-covid-codebook.csv).

## EDA Table

Each row in the data set represents COVID-19 related information for a given country on a given day starting from 2019-12-31. To limit the scope of our project, we have chosen the data set that includes information up to 2020-11-18. Within this data set, there are 57,606 observations and 50 variables. We will be primarily looking at the COVID-19 data on Canada and USA. Since we are interested in both number of new cases and number of new tests, any row with missing values in either column is removed from the date set. We also calculated the response ratio as defined below:

Response Ratio = \# of daily new tests / \# of daily new cases

Below we have shown the number of observations, mean, median and standard deviation of response ratio (Table 1).

The mean and median ratio for the USA are close to each other, which suggests that the ratio for the USA is normally distributed. The mean response ratio for Canada is higher than the median, so its distribution seems to be right skewed. This suggests that the true population of response ratio may not be normally distributed. However, since the sample size is well above 30, the large enough sample size assumption under the Central Limit Theorem is met. Therefore the sample observation does not need to come from a normally distributed population. Because all conditions of Central Limit Theorem have been met, we decided that it is safe to use mean response ratio as test statistic for our hypothesis test.

The standard deviation of Canada is quite large, compared with the mean and median ratio. It indicates that the response ratio in Canada is more widely spread out. The standard deviation of the USA is a lot smaller, which suggests that more of the ratio is clustered around the mean. It should be noted that the standard deviation might have been affected by data in the early stage of the pandemic, as tests were not fully developed and new cases were only detected within a few communities.

```{r EDA Table}
covid_all_country <-  read_csv("../data/owid-covid-data.csv", guess_max = 5000)

covid_CAN_USA <- covid_all_country %>% 
    filter(iso_code=="CAN"| iso_code =="USA") %>% 
    filter(date < "2020-11-01") %>% 
    filter(new_cases != 0 & new_tests !=0) %>% 
    select("iso_code", "date", "location", "new_tests", "new_cases") %>% 
    drop_na(new_tests, new_cases) %>% 
    mutate(response_ratio = new_tests/new_cases) 

covid_CAN_USA_summary <-  covid_CAN_USA %>% 
    group_by(iso_code) %>% 
    summarise(sample_size = n(),
              mean_response_ratio = mean(response_ratio),
              median_response_ratio = median(response_ratio),
              standard_dev = sd(response_ratio),
              )
kable(covid_CAN_USA_summary, caption = "Table 1. Summary Statistics of Response Ratio")
```

## EDA Plot

To visualize the distribution, we have plotted the distribution of each country's response ratio in the histogram below.

The histogram further confirms that the distributions of Canada's response ratio is indeed right skewed. In addition, we can see that both distributions seem to be unimodal but the centre of the distribution seem to be different. There appears to be a high concentration of the response ratio between 20 and 50 for Canada and between 10 and 20 for the USA.

```{r EDA Plot}

dist_CAN <- covid_CAN_USA %>% 
    filter(iso_code =="CAN") %>% 
    ggplot(aes(response_ratio)) +
    geom_histogram(bins=50) +
    xlab("Response Ratio, Canada") +
    ggtitle("Sample Distribution of Canada's Response Ratio")

dist_USA <- covid_CAN_USA %>% 
    filter(iso_code =="USA") %>% 
    ggplot(aes(response_ratio)) +
    geom_histogram(bins=50) +
    xlab("Response Ratio, USA") +
    ggtitle("Sample Distribution of USA's Response Ratio")

plot_grid(dist_USA, dist_CAN, ncol=1)
```

Figure 1. Histogram for Distribution of Response Ratio of Canada and USA

To better understand the entire distribution of the data, we also plotted the distribution and the mean ratio through violin plot.

```{r}
covid_CAN_USA_plot <- covid_CAN_USA %>%
  mutate(across(iso_code, fct_relabel, .fun = str_wrap, width = 30)) %>%
  ggplot(aes(x = response_ratio, y = iso_code)) +
  geom_violin(fill = "yellow", 
              alpha = .1, adjust = .8, trim = FALSE) +
  stat_summary(fun = mean,
               colour = "red",
               geom = "point") +
  scale_x_continuous(labels = scales::label_number_si()) +
  labs(title = "COVID-19 Response Ratio - Canada vs USA",
       subtitle = "Means shown in red",
       x = "Mean of Response Ratio in both countries",
       y = "Country")
covid_CAN_USA_plot
```

Figure 2. Violin Plot for Distribution of Response Ratio of Canada and USA

This is a bit hard to see given the outliers. Let's zoom into our plot to better highlight differences.

```{r}
covid_CAN_USA_plot + 
  scale_x_continuous(limits = c(0, 100))
```

Figure 3. Violin Plot for Distribution of Response Ratio of Canada and USA (Removed Outliers)

We can see that the response ratio of USA are mostly clustered around the mean while the response ratio of Canada are more widely spread.

# References
